### 2021-08-13 paper/paper.tex
Add "the" for improved reading flow

\begin{abstract}
% Abstract: at most 600 characters, written in plain English with no symbol
% nor formula + references etc.
We present Trixi.jl, a Julia package for adaptive high-order numerical simulations
of hyperbolic partial differential equations (PDEs). Utilizing Julia's strengths,
Trixi.jl is extensible, easy to use, and fast. We describe the main design choices
that enable these features and compare Trixi.jl with an established open
source Fortran code for hyperbolic PDEs using the same numerical methods.
We conclude with an assessment of Julia for simulation-focused scientific
computing, an area that is still dominated by traditional high-performance
computing languages such as C, C++, and Fortran.
\end{abstract}

@@

We are broadly interested in simulation-focused scientific computing, in particular
numerical approximations for hyperbolic partial differential equations (PDEs), computational fluid dynamics (CFD),
and related problems. The focus of our research ranges from specific applications
in CFD to general multi-physics coupling strategies to the development and analysis of high-order numerical methods
our simulations are built upon. In addition, we are academics and involved in teaching
students in these areas of science. Thus, we would like to have a
code that is
\begin{enumerate}
  \item extensible for research and development,
  \item easy to apply/understand for students/collaborators,
  \item fast enough for applied 3D problems.
\end{enumerate}
Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia} (accessed 2021-08-11)}
and wish to include all these features within a single code.
Roughly one year of collaborative work has resulted in the current version of
\trixi\footnote{\url{https://github.com/trixi-framework/Trixi.jl}}, providing adaptive
high-order numerical simulations of hyperbolic PDEs
in Julia \cite{bezanson2017julia}. Starting as an experiment, we have been able
to reach more and more of our goals with \trixi.

In this article, we present an overview of the main features and design decisions
of \trixi in Section~\ref{sec:design-of-trixi}, laying the ground for an extensible
and easy-to-use framework of high-order methods for hyperbolic PDEs. Next, we
compare the serial performance with an established high-performance computing (HPC) Fortran code in
Section~\ref{sec:performance-comparison}, demonstrating that Julia is not generically
slower than traditional HPC languages (and can even be faster in this particular case).
Thereafter, we present an assessment of Julia for simulation-focused scientific
computing based on our experience with \trixi in Section~\ref{sec:assessment-of-julia}.
Finally, we summarize our findings and conclusions in Section~\ref{sec:summary}.

@@

The design of \trixi was guided by extensibility, ease of use, and efficiency
(in this particular order). Thus, it is likely possible to make (some parts) even
faster at the cost of simplicity and readability. Nevertheless, \trixi is already quite
fast, even compared to established high-performance open source software,
as described in detail in Section~\ref{sec:performance-comparison}. In this
section, we focus mostly on the first two goals.

@@

While our Julia code is faster than the established HPC Fortran code FLUXO for this non-trivial
example, we do not claim that Julia is generally faster than Fortran, C, or C++.
Instead, we would like to emphasize that Julia code can be \emph{at least
as fast} as code written in these traditional scientific computing languages.
Further, we expect to be able to achieve similar performance with either language
by spending enough time and effort to optimize the respective code.
\trixi owes its performance optimizations in part to the code introspection and
profiling tools available in Julia. Similar tools used to optimize the performance in other
languages are often not as easy to use as their Julia counterparts.

@@

We have presented \trixi, a Julia package for adaptive high-order numerical
simulations of hyperbolic PDEs. As researchers in numerical analysis and
scientific computing, our goals were to create a framework that is extensible,
easy to use, and fast (in this particular order). Making use of Julia's strengths,
we have been successful based on recent publications making use of \trixi,
the number of students and researchers working with \trixi, and serial performance
comparisons with an established Fortran code. Having developed, from scratch, \trixi for a bit
more than one year allows us to give an assessment of Julia for simulation-focused
scientific computing.

>>>>

\begin{abstract}
% Abstract: at most 600 characters, written in plain English with no symbol
% nor formula + references etc.
We present Trixi.jl, a Julia package for adaptive high-order numerical simulations
of hyperbolic partial differential equations (PDEs). Utilizing Julia's strengths,
Trixi.jl is extensible, easy to use, and fast. We describe the main design choices
that enable these features and compare Trixi.jl with a mature open
source Fortran code for hyperbolic PDEs using the same numerical methods.
We conclude with an assessment of Julia for simulation-focused scientific
computing, an area that is still dominated by traditional high-performance
computing languages such as C, C++, and Fortran.
\end{abstract}

@@

We are broadly interested in simulation-focused scientific computing, in particular
numerical approximations for hyperbolic partial differential equations (PDEs), computational fluid dynamics (CFD),
and related problems. The focus of our research ranges from specific applications
in CFD to general multi-physics coupling strategies to the development and analysis of the high-order numerical methods
our simulations are built upon. In addition, we are academics and involved in teaching
students in these areas of science. Thus, we would like to have a
code that is
\begin{enumerate}
  \item extensible for research and development,
  \item easy to apply/understand for students/collaborators,
  \item fast enough for applied 3D problems.
\end{enumerate}
Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia} (accessed 2021-08-11)}
and wish to include all these features within a single code.
Roughly one year of collaborative work has resulted in the current version of
\trixi\footnote{\url{https://github.com/trixi-framework/Trixi.jl}}, providing adaptive
high-order numerical simulations of hyperbolic PDEs
in Julia \cite{bezanson2017julia}. Starting as an experiment, we have been able
to reach more and more of our goals with \trixi.

In this article, we present an overview of the main features and design decisions
of \trixi in Section~\ref{sec:design-of-trixi}, laying the ground for an extensible
and easy-to-use framework of high-order methods for hyperbolic PDEs. Next, we
compare the serial performance with a mature high-performance computing (HPC) Fortran code in
Section~\ref{sec:performance-comparison}, demonstrating that Julia is not generically
slower than traditional HPC languages (and can even be faster in this particular case).
Thereafter, we present an assessment of Julia for simulation-focused scientific
computing based on our experience with \trixi in Section~\ref{sec:assessment-of-julia}.
Finally, we summarize our findings and conclusions in Section~\ref{sec:summary}.

@@

The design of \trixi was guided by extensibility, ease of use, and efficiency
(in this particular order). Thus, it is likely possible to make (some parts) even
faster at the cost of simplicity and readability. Nevertheless, \trixi is already quite
fast, even compared to mature high-performance open source software,
as described in detail in Section~\ref{sec:performance-comparison}. In this
section, we focus mostly on the first two goals.

@@

While our Julia code is faster than the mature HPC Fortran code FLUXO for this non-trivial
example, we do not claim that Julia is generally faster than Fortran, C, or C++.
Instead, we would like to emphasize that Julia code can be \emph{at least
as fast} as code written in these traditional scientific computing languages.
Further, we expect to be able to achieve similar performance with either language
by spending enough time and effort to optimize the respective code.
\trixi owes its performance optimizations in part to the code introspection and
profiling tools available in Julia. Similar tools used to optimize the performance in other
languages are often not as easy to use as their Julia counterparts.

@@

We have presented \trixi, a Julia package for adaptive high-order numerical
simulations of hyperbolic PDEs. As researchers in numerical analysis and
scientific computing, our goals were to create a framework that is extensible,
easy to use, and fast (in this particular order). Making use of Julia's strengths,
we have been successful based on recent publications making use of \trixi,
the number of students and researchers working with \trixi, and serial performance
comparisons with a mature Fortran code. Having developed, from scratch, \trixi for a bit
more than one year allows us to give an assessment of Julia for simulation-focused
scientific computing.
